SIG Proceedings Paper in LaTeX Format

### 2019-09-29 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/introduction.tex
structure change w/ freq words
introduction plus abstract

This work focused on improving the convergence time of the SGNS. To facilitate our calculations we propose a slightly altered version of the original SNGS, in which we compute the loss for a very large amount of training samples, i.e 2000. The loss of such a batch was computed by taking the sum over the loss all individual training samples in the specific batch We combined this batched approach with advanced optimizers and input shuffling to decrease he convergence time of the SGNS. In combination advanced optimizers and input shuffling allowed us to decrease the convergence time of the SGNS. If combined with an optimized throughput these results could lead to an overall decrease in runtime.\\

>>>>

We propose a slightly altered version of the original SNGS, in which we compute the loss for a very large amount of training samples, i.e 2000. The loss of such a batch was computed by taking the sum over the loss all individual training samples in the specific batch. We combined this batched approach with advanced optimizers and input shuffling  This work focused on demonstrating that a batch approach, combined with advanced optimizers and input shuflling, does not hinder performance nor convergence time. If combined with an optimized throughput these results could lead to an overall decrease in runtime.\\

### 2019-09-29 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/results.tex
introduction plus abstract

\subsubsection{Adagrad}\label{ssec:results_adagrad}
Adagrad \citep{adagrad} is a very interesting optimizer for creating word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa. Words (features in our case) that appear very frequently (low learning rate) often do not have a semantic gain  that is as important as words that appear less frequently (high learning rate) to their context words. So, in theory, Adagrad is particularly well suited for our task, as for example Pennington et al. used Adagrad in the training of Glove \citep{glove}, another system which word embeddings. Our results confirmed empirically that Adagrad can work better for creating word embeddings. The  model converged in 4 epochs. When combined with shuffling Adagrad only took 3 epochs to converge, as shown in Table \ref{table:results}. These results confirms two hypothesis: first  the tendency of the Skip-Gram Model to converge faster with input shuffling and second  the big impact of having different learning rate for each feature.
Another interesting fact to notice is that a higher learning rate combined with input shuffling did not yield better results than without shuffling, in comparison to the earlier described optimzers. Adagrad performed best with a learning rate of $0.1$.
\iffalse
\begin{figure}[h]
\centering
\includegraphics[scale=0.3]{images/results_adagrad_shuffle}
\caption{Training time Adagrad with input Shuffling}
\label{fig:results_adagrad_shuffle}
\end{figure}
\fi
\subsubsection{Adam}
Adam is the most advanced of all the optimizers used in our experiments and did yield the best results as seen in Table \ref{table:results}. Adam converged in 3 epochs without shuffling and a learning rate of $0.05$ and 2 with shuffling and a learning rate of $0.001$, as shown in Table \ref{table:results}. Those are the best results achieved in any of our experiment. Exacctly as Adagrad, Adamd did not react to input shuffling the same way as SGD did.
    \begin{table}[]
\begin{tabular}{|l|l|l|}
\hline
Model    & Convergence Time & Word Similarity \\ \hline
SGD & {11}              & 0.65            \\ \hline
SGD w/shuffling & {7}              & 0.66            \\ \hline
Adam & {3}              & 0.66            \\ \hline
Adam w/ shuffling & \textbf{2}              & 0.66      \\ \hline
\end{tabular}
\label{table:results}
\caption{Results on the text8 dataset with different optimizers and input shuffling}

>>>>

\subsubsection{Adagrad}\label{ssec:results_adagrad}
Adagrad \citep{adagrad} is a very interesting optimizer for creating word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa. Words (features in our case) that appear very frequently (low learning rate) often do not have a semantic gain  that is as important as words that appear less frequently (high learning rate) to their context words. So, in theory, Adagrad is particularly well suited for our task, as for example Pennington et al. used Adagrad in the training of Glove \citep{glove}, another system which word embeddings. Our results confirmed empirically that Adagrad can work better for creating word embeddings. The  model converged in 4 epochs. When combined with shuffling Adagrad only took 3 epochs to converge, as shown in Table \ref{table:results}. These results confirms two hypothesis: first  the tendency of the Skip-Gram Model to converge faster with input shuffling and second  the big impact of having different learning rate for each feature.
Another interesting fact to notice is that a higher learning rate combined with input shuffling did not yield better results than without shuffling, in comparison to the earlier described optimzers. Adagrad performed best with a learning rate of $0.1$.
\iffalse
\begin{figure}[h]
\centering
\includegraphics[scale=0.3]{images/results_adagrad_shuffle}
\caption{Training time Adagrad with input Shuffling}
\label{fig:results_adagrad_shuffle}
\end{figure}
\fi
\subsubsection{Adam}
Adam is the most advanced of all the optimizers used in our experiments and did yield the best results as seen in Table \ref{table:results}. Adam converged in 3 epochs without shuffling and a learning rate of $0.05$ and 2 with shuffling and a learning rate of $0.001$, as shown in Table \ref{table:results}. Those are the best results achieved in any of our experiment. Exacctly as Adagrad, Adamd did not react to input shuffling the same way as SGD did.
    \begin{table}[]
\begin{tabular}{|l|l|l|}
\hline
Model    & Convergence Time & Word Similarity \\ \hline
SGD & {7}              & 0.66            \\ \hline
SGD w/shuffling & {3}              & 0.65     \\ \hline
Adam & {4}              & 0.67           \\ \hline
Adam w/ shuffling & \textbf{4}              & 0.66      \\ \hline
\end{tabular}
\label{table:results}
\caption{Results on the text8 dataset with different optimizers and input shuffling}

### 2019-09-29 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/sample-sigconf.tex
introduction plus abstract

\begin{abstract}
The Skip-Gram Model with negative sampling (SGNS) is a simple and effective algorithm to create word embeddings. The SGNS uses Stochastic Gradient Descent (SGD) as it's learning algorithm. While a lot of effort has gone into increasing the throughput of words of the SGNS, not much work has gone into optimizing the convergence time. Therefore, our work focuses on the latter. To achieve a better convergence time we use advanced optimizers and input shuffling as optimization techniques. Here are the components used to evaluate our work: to train our model we used the text8 dataset - to measure the quality of our word embeddings we used the wordsim353 dataset, which assess that similar words have vector represenations that are close to each other. We trained our model with multiple advanced optimizers: Momentum, Nesterov Accelerated Gradient, Adagrad, Adadelta, and Adam. We also applied input shuffling to the above optimizers. Adam combined with input shuffling outperformed every optimizer. Adam with shuffling also outperformed the current state of the art implementation Gensim. Adam converged to a similarity value of 0.66 (state of the art) in 2 epochs, while Gensim took 4 epochs. Hence, this work shows that advanced optimizers combined with input shuffling do decrease the convergence time of the SGNS. These results must be confirmed with other datasets, but optimizing our model in terms of throughput has the potential of further reducing the overall run time of the SGNS.

>>>>

\begin{abstract}
The Skip-Gram Model with negative sampling (SGNS) is a simple and effective algorithm to create word embeddings. The SGNS uses Stochastic Gradient Descent (SGD) as it's learning algorithm. A lot of effort has gone into increasing the throughput of words of the SGNS .  Therefore we introduce a batched approach of the SGNS that should in theory reduce the runtime of the SGNS. In our batched approach the loss is computed for a high number of training samples (i.e. 2000), which is then summed together. This sum is the used as loss for our model. This approach has the potential of further reducing the runtime of the SGNS. We combinded our batched approach with input shuffling and advanced optimizers. Here are the components used to evaluate our work: to train our model we used the text8 dataset - to measure the quality of our word embeddings we used the wordsim353 dataset, which assess that similar words have vector represenations that are close to each other. We trained our model with multiple advanced optimizers: Momentum, Nesterov Accelerated Gradient, Adagrad, Adadelta, and Adam. We also applied input shuffling to the above optimizers. Adam combined with input shuffling outperformed every optimizer. Adam with shuffling converged to a similarity value of 0.67 epoch (1\% loss of accuracy) in 4 epochs. The current state of the art implementation Gensim converges in 4 epochs to a similarity value of 0.68. Hence this work shows that our batched approach does not hinder performance and has the potential of further reducing the runtime of the SGNS. These results must be confirmed with other datasets, but optimizing our model in terms of throughput has the potential of further reducing the overall run time of the SGNS.
